---
title: "Magnitude table suppression"
author: ""
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{table suppression}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
options(rmarkdown.html_vignette.check_title = FALSE)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r include = FALSE}
htmltables <- TRUE
if (htmltables) {
    source("GaussKable.R")
    source("KableMagnitudeTable.R")
    P <- function(...) G(timevar = "geo", ...)
    M <- function(...) KableMagnitudeTable(..., numVar = "value", timevar = "geo")
} else {
    P <- function(...) cat("Formatted table not avalable")
    M <- P
}
```


## Introduction 
The `GaussSuppression` package contains several easy-to-use wrapper functions and in this vignette we will look at 
the `SuppressFewContributors` and  `SuppressDominantCells` functions.
In these functions, primary suppression is based on the number of contributors or by dominance rules.
Then, as always in this package, secondary suppression is performed using the Gauss method.

We begin by creating a dataset to be used below.

```{r}
library(GaussSuppression)
q <- data.frame(geo = rep(c("Iceland", "Portugal", "Spain"), each = 4), 
                eu = "EU", 
                sector4 = c("Agriculture", "Entertainment", "Governmental", "Industry"), 
                sector2 = "private")[-c(1, 3), ]
q$eu[1:2] <- "nonEU"
q$sector2[c(5, 9)] <- "public"
q <- q[c(2, 1, 2, 2, 3, 4, 4, 5, 5, 6, 6, 7, 8, 8, 8, 9, 9, 10), ]
q$company <- c("B", "B", "C", "D", "A", "B", "D", "B", "D", 
               "B", "D", "A", "A", "B", "C", "C", "D", "C")

q$value <- c(9.6, 16.8, 8.8, 1.9, 75.9, 7.1, 2.3, 21.6, 2, 
             25.7, 3.4, 96.6, 77.4, 11.5, 16.4, 6.5, 2.7, 8.4)
dataset <- SSBtools::SortRows(q[c(3, 4, 1, 2, 5, 6)])
rownames(dataset) <- NULL
dataset

```

We can imagine the figures in the variable `"value"` represent sales value to different sectors from different companies. 
In the first examples, we will not use the `"company"` variable, 
but instead assume that each row represents the contribution of a unique company.
Our input data can then be illustrated like this:
\
```{r echo=FALSE}
M(caption = '**Table 1**: Input data with the 18 "value" contributions.',
  dataset, formula = ~sector4:geo-1)
```
\

## Initial basic examples

### Using few contributors primary suppression

In the first example, we use `SuppressFewContributors` with `maxN = 2`. 
This means that aggregates from one and two contributors are primary suppressed.

```{r}
SuppressFewContributors(data=dataset, 
                        numVar = "value", 
                        dimVar= c("sector4", "geo"), 
                        maxN=2)
```

In the output, the number of contributors is in columns `nRule` and `nAll`. 
The two columns are equal under normal usage.

A formatted version of this output is given in Table 2 below.
Primary suppressed cells are underlined and labeled in red, while the secondary suppressed cells are labeled in purple.
\
```{r echo=FALSE}
P(caption = '**Table 2**: Output from `SuppressFewContributors` with `maxN = 2` (number contributors in parenthesis)',
  data=dataset, 
  numVar = "value", 
  dimVar= c("sector4", "geo"), 
  maxN=2,
  fun = SuppressFewContributors,
  print_expr = 'paste0(value, " (",nAll ,") ")')
```
\

### Using dominant cell primary suppression

In the second example, we use `SuppressDominantCells`  with `n = 2` and `k = 90`.
This means that aggregates are primary suppressed whenever the 
two largest contributions exceeds 90% of the cell total.

```{r}
SuppressDominantCells(data=dataset, 
                      numVar = "value", 
                      dimVar= c("sector4", "geo"), 
                      n = 2, k = 90, allDominance = TRUE)
```

To see the percentage from the two largest contributions, `allDominance = TRUE` was used.
A formatted version of this output is given in Table 3 below.
\
```{r echo=FALSE}
P(caption = '**Table 3**: Output from `SuppressDominantCells`  with `n = 2` and `k = 90` (percentage from two largest contributions in parenthesis)',
  data=dataset, 
  numVar = "value", 
  dimVar= c("sector4", "geo"), 
  n=2, k=90, allDominance = TRUE,
  fun = SuppressDominantCells,
  print_expr = 'paste0(value, " (",round(100*`primary.2:90`) ,"%) ")')
```
\
\

## An hierarchical table and two dominance rules  

Here we use `SuppressDominantCells`  with `n = 1:2` and `k = c(80, 90)`.
This means that aggregates are primary suppressed whenever 
the largest contribution exceeds 80% of the cell total 
or when the two largest contributions exceed 90% of the cell total.

In addition, the example below is made even more advanced by including the variables "sector2" and "eu". 

```{r}
output <- SuppressDominantCells(data=dataset, 
                                numVar = "value", 
                                dimVar= c("sector4", "sector2", "geo", "eu"), 
                                n = 1:2, k = c(80, 90))
head(output)
```

Formatted version:
\
```{r echo=FALSE}
P(caption = '**Table 4**: Output from `SuppressDominantCells`  with `n = 1:2` and `k = c(80, 90)`',
  data=dataset, 
  numVar = "value", 
  dimVar= c("sector4", "sector2", "geo", "eu"), 
  n = 1:2, k = c(80, 90), 
  fun = SuppressDominantCells,
  print_expr = 'value')
```
\

As described in the define-tables vignette hierarchies are here treated automatically.
The same output is obtained if we first generate hierarchies by: 

```{r}
dimlists <- SSBtools::FindDimLists(dataset[c("sector4", "sector2", "geo", "eu")])
dimlists
```

And thereafter run SuppressDominantCells with these hierarchies as input:

```{r}
output <- SuppressDominantCells(data=dataset, 
                                numVar = "value", 
                                hierarchies = dimlists,  
                                n = 1:2, k = c(80, 90))
```
\
\

## Using the formula interface

Using the formula interface is one way to achieve fewer cells in the output:

```{r}
output <- SuppressFewContributors(data=dataset, 
                                  numVar = "value", 
                                  formula = ~sector2*geo + sector4*eu, 
                                  maxN=2,
                                  removeEmpty = FALSE)
head(output)
tail(output)
```

In the formatted version of this output, blank cells indicate that they are not included in the output. 
\
```{r echo=FALSE}
P(caption = '**Table 5**: Output from `SuppressFewContributors` with `maxN = 2` (number contributors in parenthesis)',
                                  data=dataset, 
                                  numVar = "value", 
                                  formula = ~sector2*geo + sector4*eu, 
                                  maxN=2,
                                  removeEmpty = FALSE,
                                  fun = SuppressFewContributors,
                                  print_expr = 'paste0(value, " (",nAll ,") ")')
```
\

Please note that in order to include the three empty cells with no contributors, 
the `removeEmpty` parameter was set to `FALSE`. By default, this parameter is set to `TRUE` when using the formula interface. 
\
\

## Using the contributorVar parameter (`contributorVar = "company"`)

According to the `"company"` variable in the data set, 
there are only four contribution companies (A, B, C and D).
When this is taken into account, the primary suppression rules will be applied to data that, 
within each cell, is aggregated within each contributor.
Our example data aggregated in this way is shown below.
\
```{r echo=FALSE}
M(caption = '**Table 6**: The "value" data aggregated according to hierarchy and contributor',
  dataset, dimVar = c("sector4", "sector2", "geo", "eu"),   contributorVar = "company")
```
\

Below we take into account contributor IDs  when using few contributors primary suppression.

```{r}
output <- SuppressFewContributors(data=dataset, 
                                  numVar = "value", 
                                  dimVar = c("sector4", "sector2", "geo", "eu"),
                                  maxN=2,
                                  contributorVar = "company")
head(output)
```

Formatted version: 
\
```{r echo=FALSE}
P(caption = '**Table 7**: Output from `SuppressFewContributors` with `maxN = 2` and with `contributorVar = "company"` (number contributors in parenthesis)',
                                  data=dataset, 
                                  numVar = "value", 
                                  dimVar = c("sector4", "sector2", "geo", "eu"), 
                                  maxN=2,
                                  contributorVar = "company",
                                  fun = SuppressFewContributors,
                                  print_expr = 'paste0(value, " (",nAll ,") ")')
```
\

Below we take into account contributor IDs when using dominant cell primary suppression.

```{r}
output <- SuppressDominantCells(data=dataset, 
                                numVar = "value", 
                                formula = ~sector2*geo + sector4*eu,
                                contributorVar = "company",
                                n = 1:2, k = c(80, 95))
head(output)
```

Formatted version: 
\
```{r echo=FALSE}
P(caption = '**Table 8**: Output from `SuppressDominantCells`  with `n = 1:2` and `k = c(80, 95)` and with `contributorVar = "company"`',
  data=dataset, 
  numVar = "value", 
  formula = ~sector2*geo + sector4*eu,
  contributorVar = "company",
  n = 1:2, k = c(80, 95), 
  fun = SuppressDominantCells,
  print_expr = 'value')
```
\
\

## The singleton problem 


Below, the data is suppressed in the same way as in Table 7, but with a different formula.
\

```{r}
output <- SuppressDominantCells(data=dataset,
                                numVar = "value", 
                                formula = ~sector4*geo + sector2*eu,
                                contributorVar = "company",
                                n = 1:2, k = c(80, 95))
head(output)
```
\

Formatted version:
\
```{r echo=FALSE}
P(caption = '**Table 9**: Output from `SuppressDominantCells`  with `n = 1:2` and `k = c(80, 95)` and with `contributorVar = "company"`',
  data=dataset, 
  numVar = "value", 
  formula = ~sector4*geo + sector2*eu,
  contributorVar = "company",
  n = 1:2, k = c(80, 95), 
  fun = SuppressDominantCells,
  print_expr = 'value')
```
\

By using `singletonMethod = "none"` in this case, *Entertainment:Spain* will not be suppressed. 
This cell is suppressed due to the default handling of the singleton problem. 
The reason is that *Iceland:Spain* has a single contributor. This contributor can reveal *Portugal:Spain* if *Entertainment:Spain* is not suppressed.

