---
title: "Small count frequency table suppression"
author: ""
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Small count frequency table suppression}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
options(rmarkdown.html_vignette.check_title = FALSE)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r include = FALSE}
htmltables = TRUE
if(htmltables){
  source("GaussKable.R")
  P = function(...) G(fun = SuppressSmallCounts, timevar = "geo", ...)
} else { 
  P = function(...) cat("Formatted table not avalable")
}
```


## Introduction 
The `GaussSuppression` package contains several easy-to-use wrapper functions and in this vignette we will look at 
the `SuppressSmallCounts` function.
In this function, small frequencies are primary suppressed.
Then, as always in this package, secondary suppression is performed using the Gauss method.

We begin by creating datasets to be used below.
The first examples are based on `dataset_a`, which has six rows.

```{r}
library(GaussSuppression)
dataset <- SSBtoolsData("sprt_emp_withEU")[, c(1, 2, 5, 3)]
dataset$age[dataset$age == "Y15-29"] <- "young"
dataset$age[dataset$age == "Y30-64"] <- "old"
dataset$freq <- c(5, 2, 0, 6, 3, 4, 5, 0, 0, 6, 3, 4, 7, 1, 1, 5, 4, 3)
dataset_a <- dataset[dataset$year == "2014", -4]
dataset_b <- dataset[dataset$year == "2015", -4]
dataset_a
```

## An initial basic example

In the function description (`?SuppressSmallCounts`), the only visible parameter is `maxN` in addition to the parameters considered in the define-tables vignette.
In the first example, we use `maxN = 1` which means that zeros and ones are primary suppressed.

```{r}
SuppressSmallCounts(data = dataset_a, 
                    dimVar = c("age", "geo"), 
                    freqVar = "freq", 
                    maxN = 1)
```

A formatted version of this output is given in Table 1 below. Primary suppressed cells are underlined and labeled in red, while the secondary suppressed cells are labeled in purple.

```{r echo=FALSE}
P(caption = "Table 1", data = dataset_a, 
                    dimVar = c("age", "geo"), 
                    freqVar = "freq", 
                    maxN = 1)
```

The same output is obtained if microdata is sent as input as illustrated by de code below.  

```{r}
microdata_a <- SSBtools::MakeMicro(dataset_a, "freq")[-4]
output <- SuppressSmallCounts(data = microdata_a, 
                              dimVar = c("age", "geo"), 
                              maxN = 1)
```

A related point is that the third row of the table can be omitted (`data = dataset_a[-3, ]`) since the frequency is zero. 
Later in this vignette, we address scenarios where the inclusion of zeros may be important.

##  An hierarchical table

A more advanced example is obtained by including the variable "eu". 

```{r}
SuppressSmallCounts(data = dataset_a, 
                    dimVar = c("age", "geo", "eu"), 
                    freqVar = "freq", 
                    maxN = 2)
```

A formatted version of this output: 

```{r echo=FALSE}
P(caption = "Table 2", data = dataset_a, 
                    dimVar = c("age", "geo", "eu"), 
                    freqVar = "freq", 
                    maxN = 2)
```

As described in the define-tables vignette hierarchies are here treated automatically.
The same output is obtained if we first generate hierarchies by: 

```{r}
dimlists <- SSBtools::FindDimLists(dataset_a[c("age", "geo", "eu")])
dimlists
```

And thereafter run SuppressSmallCounts with these hierarchies as input:

```{r eval=FALSE}
SuppressSmallCounts(data = dataset_a[c("age", "geo", "freq")], 
                    hierarchies = dimlists, 
                    freqVar = "freq", 
                    maxN = 2)
```


## Using the formula interface

Using the formula interface is one way to achieve fewer cells in the output: 


```{r}
SuppressSmallCounts(data = dataset_a, 
                    formula = ~age:eu + geo, 
                    freqVar = "freq", 
                    maxN = 2)
```

In the formatted version of this output, blank cells indicate that they are not included in the output. 

```{r echo=FALSE}
P(caption = "Table 3", data = dataset_a, 
                    formula = ~age:eu + geo, 
                    freqVar = "freq", 
                    maxN = 2)
```


## About suppression of zeros

By default, zeros are suppressed in order to protect against attribute disclosure in frequency tables. 
However, there are exceptions.  

### Zeros not suppressed

One option is to use `protectZeros = FALSE`. 

```{r}
SuppressSmallCounts(data = dataset_a, 
                    dimVar = c("age", "geo", "eu"),  
                    freqVar = "freq", 
                    maxN = 4,
                    protectZeros = FALSE)
```

Formatted version:

```{r echo=FALSE}
P(caption = "Table 4", data = dataset_a, 
                    dimVar = c("age", "geo", "eu"),  
                    freqVar = "freq", 
                    maxN = 4,
                    protectZeros = FALSE)
```

Another possibility that gives the same output is:

```{r}
output <- SuppressSmallCounts(data = dataset_a[-3, ], 
                              dimVar = c("age", "geo", "eu"),  
                              freqVar = "freq", 
                              maxN = 4,
                              extend0 = FALSE, 
                              structuralEmpty = TRUE)
```

Here the zero-frequency row is omitted in input. 
By default, the table is automatically extended so that the Gauss algorithm handles zeros correctly. 
When this is turned off (`extend0 = FALSE`), a warning with the following text will appear: 
"*Suppressed cells with empty input will not be protected. Extend input data with zeros?*".
However, with `structuralEmpty = TRUE`, the "empty zeros" are assumed to represent structural zeros 
that must not be suppressed. 
As exemplified a little further below, one can thus handle data with both structural and non-structural zeros.


### Secondary suppressed zeros 

We can combine `protectZeros = FALSE` with `secondaryZeros = TRUE`. 

```{r}
SuppressSmallCounts(data = dataset_a, 
                    dimVar = c("age", "geo", "eu"),  
                    freqVar = "freq", 
                    maxN = 3,
                    protectZeros = FALSE, 
                    secondaryZeros = TRUE)
```

Formatted version:

```{r echo=FALSE}
P(caption = "Table 5", data = dataset_a, 
                    dimVar = c("age", "geo", "eu"),  
                    freqVar = "freq", 
                    maxN = 3,
                    protectZeros = FALSE, 
                    secondaryZeros = TRUE)
```

### Both structural and non-structural zeros

The example below uses `dataset_b`, which has two zeros.

```{r}
  dataset_b
```

Now we omit one of the zeros and keep the other.
In general, we will omit all structural zeros.

```{r}
SuppressSmallCounts(data = dataset_b[-2,  ], 
                    dimVar = c("age", "geo", "eu"),  
                    freqVar = "freq", 
                    maxN = 2,
                    extend0 = FALSE, 
                    structuralEmpty = TRUE)
```

Formatted version:

```{r echo=FALSE}
P(caption = "Table 6", data = dataset_b[-2,  ], 
                    dimVar = c("age", "geo", "eu"),  
                    freqVar = "freq", 
                    maxN = 2,
                    extend0 = FALSE, 
                    structuralEmpty = TRUE)
```

Now, the data has been processed correctly, the structural zeros will be published while the other zeros are suppressed.

To get the same output with the formula interface, we can use the following code:

```{r eval = FALSE}
SuppressSmallCounts(data = dataset_b[-2,  ], 
                    formula = ~age * (geo + eu),  
                    freqVar = "freq", 
                    maxN = 2,
                    extend0 = FALSE, 
                    structuralEmpty = TRUE,
                    removeEmpty = FALSE)
```

Please note that in order to include empty cells in the output, you need to set the `removeEmpty` parameter to `FALSE`. 
By default, this parameter is set to `TRUE` when using the formula interface.


## The problem of singletons and zeros

### The problem of zeros

When using the standard suppression technique on table `dataset_b`, many cells are suppressed. 

```{r}
SuppressSmallCounts(data = dataset_b, 
                    dimVar = c("age", "geo", "eu"),  
                    freqVar = "freq", 
                    maxN = 2)
```

Formatted version:

```{r echo=FALSE}
P(caption = "Table 7", data = dataset_b, 
                    dimVar = c("age", "geo", "eu"),  
                    freqVar = "freq", 
                    maxN = 2)
```


The reason is that it is easy to reveal the zeros if *young:Spain* is not suppressed. 
Then the sum of *young:Iceland* and *young:Portugal* can easily be calculated to be zero. 
Since negative frequencies are not possible, the only possibility is two zeros. 

The handling of this problem is standard, but it can be turned off by `singletonMethod = "none"`.


### The problem of singletons

This problem occurs when `protectZeros = FALSE` and `secondaryZeros = FALSE` (default).
We now also look at a larger example that uses `dataset` which has 18 rows.   

```{r}
output <- SuppressSmallCounts(data = dataset, 
                              formula = ~age*geo*year + eu*year,  
                              freqVar = "freq", 
                              maxN = 1, 
                              protectZeros = FALSE)
head(output)
```

Formatted version:

```{r echo=FALSE}
P(caption = "Table 8", data = dataset, 
                    formula = ~age*geo*year + eu*year,
                    freqVar = "freq", 
                    maxN = 1,
                    protectZeros = FALSE)
```

By using `singletonMethod = "none"` in this case, *young:2016:Spain* will not be suppressed.
Then the sum of *young:2016:Iceland* and *young:2016:Portugal* can easily be calculated to be two.
Since zeros are never suppressed, the only possibility is two ones. 

However, in our output, *young:2016:Spain* is suppressed due to the standard handling of this problem.

